---
title: "Day 6 Homework"
output: html_document
---

0. Get data ready.

```{r}

library(reshape2) 
library(ggplot2)

mydata <- read.csv("EEG_compiled.txt")
mydata$subject <- sub("^", "S", mydata$subject) # add "S" in front of the subject number in column "subj"
head(mydata) # check that it worked

# make subj, cond, elec factors
mydata$subject <- factor(mydata$subject)
mydata$condition <- factor(mydata$condition)
mydata$electrode <- factor(mydata$electrode)
summary(mydata) # check that it worked

# relabel levels of factors
levels(mydata$condition) <- c("definite", "indefinite")
summary(mydata) # check that it worked
levels(mydata$electrode) <- c("O2", "O1", "FP2", "Pz", "P4", "CP4", "P8", "C4", "TP8", "T8", "P7", "P3", "CP3", "CPz", "Cz", "FC4", "FT8", "TP7", "C3", "FCz", "Fz", "F4", "F8", "T7", "FT7", "FC3", "F3", "elec28", "F7", "FP1", "elec31", "elec32")
summary(mydata$electrode) # check that it worked
summary(mydata)



```


1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  

2. Write out (in plain English, in the notebook) what you are trying to do

For each of the subjects, I would like to take the mean amplitude for each electrode for each condition in the timewindow of 300 to 500 ms, which corresponds to timesteps 200 to 300.
Then I would like to do a t-test to see if there is a significant difference between conditions.

3. Write some "pseudo-code" comments in an R chunk, as an outline

```{r}

# create vector containing all the levels of the "subjects" factor

# create container for answers, with one space for each subject to put the t-test results

#start loop
for...# go through this loop for each member of the subject vector 
  # take the subset of the data for the current subject
  # take a subset of the data that includes only timepoints 200 to 300
  # aggregate over timepoints, take the mean of "amplitude" for each electrode for each condition
  # assign values in column "definite" to one variable, and values in column "indefinite" to another       variable
  # do a t-test on those variables
  # assign t-test result to container



```

4. Write some (real) code that will do what you want ONCE, on one subset of the data.

```{r}

# take a subset of the data for one participant
subset1 <- mydata[mydata$subject %in% "S2", ]

# take a subset of the data that includes only timepoints 200 to 300
subset2 <- subset1[subset1$timestep %in% c(200:300), ]
                                      
# aggregate over timepoints, take the mean of "amplitude" for each electrode for each condition
data.agg <- dcast(subset2, electrode ~ condition, value.var = "amplitude", fun.aggregate = mean)
head(data.agg)

# do a t-test
definite <- data.agg$definite
indefinite <- data.agg$indefinite

results <- t.test(definite, indefinite, paired = TRUE)
print(results)

```



5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.

```{r}

summary(mydata$subject)

# create vector with subject numbers
subjects <- c("S11", "S13", "S16", "S18", "S2", "S20", "S24", "S25", "S28", "S29", "S3", "S30", "S31", "S32", "S33", "S34", "S35", "S36", "S37", "S38", "S39", "S40", "S42", "S43", "S44", "S48", "S49", "S5", "S7", "S8", "S9")


rm(my.answer) 
my.answer <- vector(length = length(subjects)) #create container for answers
for(this.counter in subjects) {
  this.subject <- subjects[this.counter]
  subset1 <- mydata[mydata$subject %in% this.subject, ]
  subset2 <- subset1[subset1$timestep %in% c(200:300), ]
  data.agg <- dcast(subset2, electrode ~ condition, value.var = "amplitude", fun.aggregate = mean)
  definite <- data.agg$definite
  indefinite <- data.agg$indefinite
  results <- t.test(definite, indefinite, paired = TRUE)
  my.answer[this.counter] <- results
}

# I can't get this to work, so I'll try doing it the slightlier easier way:

my.answer.t_value <- data.frame(length = 32) 
my.answer.df <- data.frame(length = 32) 
my.answer.p_value <- data.frame(length = 32) 
for(this.subject in c("S11", "S13", "S16", "S18", "S2", "S20", "S24", "S25", "S28", "S29", "S3", "S30", "S31", "S32", "S33", "S34", "S35", "S36", "S37", "S38", "S39", "S40", "S42", "S43", "S44", "S48", "S49", "S5", "S7", "S8", "S9")) {
  subset1 <- mydata[mydata$subject %in% this.subject, ]
  subset2 <- subset1[subset1$timestep %in% c(200:300), ]
  data.agg <- dcast(subset2, electrode ~ condition, value.var = "amplitude", fun.aggregate = mean)
  definite <- data.agg$definite
  indefinite <- data.agg$indefinite
  results <- t.test(definite, indefinite, paired = TRUE)
  my.answer.t_value[this.subject] <- results$statistic
  my.answer.df[this.subject] <- results$parameter
  my.answer.p_value[this.subject] <- results$p.value
}

# combine t-value, df, and p-value for all subjects into one dataframe
my.answer <- rbind(my.answer.t_value, my.answer.df, my.answer.p_value)
print(my.answer)

```



6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.

Remove data from subject 38

```{r}
# clear objects
rm(my.answer.t_value)
rm(my.answer.df)
rm(my.answer.p_value)
rm(my.answer)

# create containers
my.answer.t_value <- data.frame(length = 32) 
my.answer.df <- data.frame(length = 32) 
my.answer.p_value <- data.frame(length = 32) 
for(this.subject in c("S11", "S13", "S16", "S18", "S2", "S20", "S24", "S25", "S28", "S29", "S3", "S30", "S31", "S32", "S33", "S34", "S35", "S36", "S37", "S38", "S39", "S40", "S42", "S43", "S44", "S48", "S49", "S5", "S7", "S8", "S9")) {
  if (!this.subject == "S38") {
  subset1 <- mydata[mydata$subject %in% this.subject, ]
  subset2 <- subset1[subset1$timestep %in% c(200:300), ]
  data.agg <- dcast(subset2, electrode ~ condition, value.var = "amplitude", fun.aggregate = mean)
  definite <- data.agg$definite
  indefinite <- data.agg$indefinite
  results <- t.test(definite, indefinite, paired = TRUE)
  my.answer.t_value[this.subject] <- results$statistic
  my.answer.df[this.subject] <- results$parameter
  my.answer.p_value[this.subject] <- results$p.value
  }
}

# combine t-value, df, and p-value for all subjects into one dataframe
my.answer <- rbind(my.answer.t_value, my.answer.df, my.answer.p_value)
print(my.answer)




```





