---
title: "Day 3 Homework"
output: html_document
---

The goal of this homework is to get practice with manipulating data.

0. Get everything ready

```{r}

library(reshape2) # load reshape2 package

mydata <- read.csv("EEG_compiled.txt")
summary(mydata)
head(mydata)

mydata$subject <- sub("^", "S", mydata$subject) # add "S" in front of the subject number in column "subj"
head(mydata) # check that it worked

# make subj, cond, elec factors
mydata$subject <- factor(mydata$subject)
mydata$condition <- factor(mydata$condition)
mydata$electrode <- factor(mydata$electrode)
summary(mydata) # check that it worked

# relabel levels of factors
levels(mydata$condition) <- c("definite", "indefinite")
summary(mydata) # check that it worked
levels(mydata$electrode) <- c("O2", "O1", "FP2", "Pz", "P4", "CP4", "P8", "C4", "TP8", "T8", "P7", "P3", "CP3", "CPz", "Cz", "FC4", "FT8", "TP7", "C3", "FCz", "Fz", "F4", "F8", "T7", "FT7", "FC3", "F3", "elec28", "F7", "FP1", "elec31", "elec32")
summary(mydata$electrode) # check that it worked
summary(mydata)

```



1. Make both a "wide" and a "long" version of your data
 

```{r}

# make a wide version of the data
mydata.wide <- mydata 
mydata.wide <- dcast(mydata, subject + timestep ~ condition + electrode, value.var = "amplitude") 
head(mydata.wide)


# make a long version of the data
mydata.long <- mydata.wide
mydata.long <- melt(mydata.wide, id.vars = c("subject", "timestep"), value.name = "amplitude", variable.name = "condition.electrode")
head(mydata.long)


```

```{r eval=FALSE}
# casting with a lot of variables

mydata.wide <- dcast(mydata.long, [80 columns here] ~ split)

columns.to.keep <- colnames(mydata.long)[1:80]
columns.tokeep.pasted <- paste(columns.to.keep, callapse = " + ") # takes a vector and pastes together all the things in the vector, so this creates one string!
my.cast.formula <- as.formula(paste(columns.to.keep.pasted, " ~ split"))
class(my.cast.formula) # should be of type formula
mydata.wide <- dcast(mydata.long, formula = my.cast.formula)


mydata.wide <- dcast(mydata.long, [ ,c(1,2,4)] ~ condition.electrode)

columns.to.keep <- colnames(mydata.long[ , c(1,2,4)])
columns.to.keep.pasted <- paste(columns.to.keep, callapse = " + ") # takes a vector and pastes together all the things in the vector, so this creates one string!
my.cast.formula <- as.formula(paste(columns.to.keep.pasted, " ~ condition.electrode")) # says 'object 'subject' not found'
class(my.cast.formula) # should be of type formula
mydata.wide <- dcast(mydata.long, formula = my.cast.formula)


```






2. Transform a variable

```{r}

mydata.ms <- mydata
summary(mydata.ms) # check that data is in the format I think it is

# create a column containing the timesteps in milliseconds (each timestep is 2ms)
mydata.ms$ms <- mydata.ms$timestep*2
head(mydata.ms)
summary(mydata.ms)

```



3. Recode a factor

```{r}

# I am not 100% sure what you mean by "recode" here, but since I relabeled factors at the beginning, I'm reordering levels 

mydata.indef.def <- mydata
summary(mydata.indef.def) 

# reorder the levels of factor "condition"
mydata.indef.def$condition <- factor(mydata.indef.def$condition, levels = c("indefinite", "definite"))
levels(mydata.indef.def$condition)
summary(mydata.indef.def) # check that it worked

```



4. Add a new factor, where the levels are conditional on the values of some other variable. E.g., a factor of "high", "medium", and "low" based on thresholds for another variable.

```{r}

# adding factor "scalp", with levels "anterior" and "posterior", "other"
# anterior electrodes: F7, FT7, FP1, F3, FC3, Fz, FCz, FP2, F4, FC4, F8, FT8
# posterior electrodes: TP7, P7, CP3, P3, O1, CPz, Pz, CP4, P4, O2, TP8, P8

mydata.scalp <- mydata

mydata.scalp$scalp <- "other"
head(mydata.scalp)
mydata.scalp$scalp[mydata.scalp$electrode %in% c("F7", "FT7", "FP1", "F3", "FC3", "Fz", "FCz", "FP2", "F4", "FC4", "F8", "FT8")] <- "anterior"
mydata.scalp$scalp[mydata.scalp$electrode %in% c("P7", "TP7", "O1", "P3", "CP3", "Pz", "CPz", "O2", "P4", "CP4", "P8", "TP8")] <- "posterior"
xtabs(~ scalp + electrode, data = mydata.scalp) # cross-tabulation to check if it worked

summary(mydata.scalp) # shows that "scalp" is not a factor 


# I don't understand this following part, is it an alternative way of doing the same thing?
# mydata.scalp2 <- cbind(mydata.scalp, 
#                       colsplit(mydata.scalp$scalp, pattern = "\\.",
#                                names = c("scalp")))
# myiris.melt2 <- cbind(myiris.melt, 
#                       colsplit(myiris.melt$measurement.type, pattern = "\\.",
#                                names = c("PlantPart", "Dimension")))



```

Class notes
```{r}



```




5. Reorder the levels of this new factor

```{r}

summary(mydata.scalp) # shows that "scalp" is not a factor here!

# mydata.scalp$scalp <- factor(mydata.scalp$scalp) # turns "scalp" into a factor, not really necessary
# levels(mydata.scalp$scalp) # check levels of factor "scalp"

# reorder the levels of factor "scalp"
mydata.scalp$scalp <- factor(mydata.scalp$scalp, levels = c("other", "anterior", "posterior"))
levels(mydata.scalp$scalp) # check levels of factor "scalp"

```


6. Choose a factor that could be re-coded as multiple dimensions (recall the sepal/petal vs. length/width breakdown). If you don't have one, try to create one.  Now create two (or more) new factors that represent these dimensions.

```{r}
# create factor "quadrant", with 5 levels
# other
# front.left: "FP1", "F3", "FC3", "F7", "FT7"
# front.right: "FP2", "F4", "FC4", "F8", "FT8"
# back.left: "O1", "P3", "CP3", "P7", "TP7"
# back.right: "O2", "P4", "CP4", "P8", "TP8"

summary(mydata) # make sure the data is in the format I think it is
mydata.quad <- mydata

mydata.quad$quadrant <- "other" # create column "quadrant" and fill all rows with "other"
head(mydata.quad)
# change value in "quadrant" to the appropriate value for these electrodes
mydata.quad$quadrant[mydata.quad$electrode %in% c("FP1", "F3", "FC3", "F7", "FT7")] <- "front.left" 
mydata.quad$quadrant[mydata.quad$electrode %in% c("FP2", "F4", "FC4", "F8", "FT8")] <- "front.right"
mydata.quad$quadrant[mydata.quad$electrode %in% c("O1", "P3", "CP3", "P7", "TP7")] <- "back.left"
mydata.quad$quadrant[mydata.quad$electrode %in% c("O2", "P4", "CP4", "P8", "TP8")] <- "back.right"
xtabs(~ quadrant + electrode, data = mydata.quad) # cross-tabulation to check if it worked

summary(mydata.quad) # shows that "quadrant" is not a factor
mydata.quad$quadrant <- factor(mydata.quad$quadrant) # make "quadrant" a factor
summary(mydata.quad)

#split along the left/right dimension

# delete "other" electrodes by making new dataframe
mydata.quad2 <- mydata.quad[mydata.quad$quadrant %in% c("front.left", "front.right", "back.left", "back.right"), ]
# mydata.quad2 <- mydata.quad[!mydata.quad$quadrant %in% c("other"), ] # this would do the same

summary(mydata.quad2)
mydata.quad2 <- droplevels(mydata.quad2) # remove level "other" from factor "quadrant"
summary(mydata.quad2)

mydata.cast <- dcast(mydata.quad2, subject + condition + electrode + timestep ~ quadrant, value.var = "amplitude" ) # split "quadrant" into 4 columns by level
head(mydata.cast)



# When I run this line, R freezes and I have to restart it. Why?
# mydata.quad3 <- cbind(mydata.quad2, colsplit(mydata.quad2$quadrant, pattern = "\\.", names = c("hemisphere", "front.back")))

# fixing this
# I don't think this one was necessary: new.cols <- colsplit(as.character(mydata.quad2$quadrant), pattern"\\.", names = c("hemisphere", "front.back"))

mydata.quad2$hemisphere <- ifelse(grepl("left", mydata.quad2$quadrant), "left", "right") #grepl secahres for "left" in the dataset and returns a logical TRUE or FALSE, which is then replaced by "left" and "right"
mydata.quad2$front.back <- ifelse(grepl("back", mydata.quad2$quadrant), "back", "front")
xtabs(~ hemisphere + front.back, data = mydata.quad2)








```



7. Using `dcast`, aggregate some variable into "cell means" of some kind.

```{r}
# this line breaks R, it triggers a message saying that R encountered a fatal error and has to relaunch. Why?
#data.agg <- dcast(mydata.quad2, subject + condition + timestep + electrode ~ quadrant, value.var = "amplitude", fun.aggregate = mean)

# make a smaller dataset
data.small <- mydata[mydata$timestep == 1, ] # make a new dataset for the subset where "timestep"" is 1
data.small$timestep <- NULL # delete "timestep" column
summary(data.small)

# take the mean of "amplitude" for each electrode for each condition (across subjects)
data.agg <- dcast(data.small, condition ~ electrode, value.var = "amplitude", fun.aggregate = mean)
data.agg

# take the mean of "amplitude" for each condition for each electrode (across subjects)
data.agg2 <- dcast(data.small, electrode ~ condition, value.var = "amplitude", fun.aggregate = mean)
head(data.agg2)

# take the mean of "amplitude" for each condition for each subject (across electrodes)
data.agg3 <- dcast(data.small, subject ~ condition, value.var = "amplitude", fun.aggregate = mean)
data.agg3
  

```


8.  Save all of the above in a well-documented R Markdown document.  Email me the .Rmd file (and if you use different or updated data from Day 2 Homework, send that to me, too).
